\large{\bf Looking forwards and backwards: dynamics and genealogies of locally regulated populations }

### 2023-06-15 forwards-and-backwards.tex
EJP format

\documentclass[12pt]{article}
\usepackage{fullpage}
\usepackage{amssymb}
\usepackage{ntheorem}
\usepackage{graphics} 
\usepackage{amsmath}
\usepackage[color=yellow]{todonotes}
\usepackage{url}
\usepackage{xcolor}
\usepackage{longtable}
\usepackage[hidelinks]{hyperref}
\usepackage{bbm}

\usepackage{natbib}

\newenvironment {proof}{{\noindent\bf Proof }}{\hfill $\Box$ \medskip}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{condition}[theorem]{Condition}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{hypothesis}[theorem]{Hypothesis}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{example}[theorem]{Example}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{assumptions}[theorem]{Assumptions}
\newtheorem{terminology}[theorem]{Terminology}
\newtheorem*{terminology-non}[theorem]{Terminology}

\renewcommand{\theequation}{\arabic{section}.\arabic{equation}}
\def \non{{\nonumber}}
\def \hat{\widehat}
\def \tilde{\widetilde}
\def \bar{\overline}
\newcommand{\IP}{\mathbb P}
\newcommand{\IQ}{\mathbb Q}
\newcommand{\IE}{\mathbb E}
\newcommand{\IR}{\mathbb R}
\newcommand{\IZ}{\mathbb Z}
\newcommand{\IN}{\mathbb N}
\newcommand{\IT}{\mathbb T}
\newcommand{\IC}{\mathbb C}
\newcommand{\ind}{\mathbf{1}}
\newcommand{\bigO}{\mathcal{O}}
\newcommand{\grad}{\nabla}
\newcommand{\dif}{\mathrm{d}\,}

%%%%%%% notation
\newcommand{\DG}{\mathcal{B}}  % generator of the dispersal process
\newcommand{\DD}{\mathcal{D}}  % the second-order part of the generator of dispersal
\newcommand{\meanq}{\vec b}    % mean of dispersal, times theta
\newcommand{\covq}{\mathbf{C}}     % covariance matrix of dispersal, times theta
\newcommand{\kernel}{\rho}  % interaction kernels
\newcommand{\smooth}[1]{\kernel_{#1} \! * \!}  % convolution by the interaction kernel
\newcommand{\wavespeed}{\mathfrak{c}}    % speed (vector) of a wave
\newcommand{\Lgen}{\mathcal{L}}    % generator of a lineage
\newcommand{\Pgen}{\mathcal{P}}    % generator of the population process
\newcommand{\lp}{\xi}              % process with levels
\newcommand{\labelspace}{\mathcal{I}} % space of labels
\newcommand{\concat}{\oplus}   % concatenation of labels
\newcommand{\measures}{\mathcal{M}_F(\IR^d)} % finite measures on Rd
\newcommand{\cmeasures}{\mathcal{M}_F(\overline{\IR}^d)} % finite measures on compactified Rd
\newcommand{\lpmeasures}{\mathcal{M}(\overline{\IR^d} \times [0,\infty))} % locally finite measures on (compactified space) x level
\newcommand{\flpmeasures}{\mathcal{M}(\IR^d \times [0,\infty))} % locally finite measures on space x level

\newcommand{\plr}[1]{\todo[inline]{Peter: #1}}
\newcommand{\comment}[1]{{\color{blue} \it #1}}

\numberwithin{equation}{section}

@@

\title{\large{\bf
Looking forwards and backwards: dynamics and genealogies of locally regulated populations
}}

\author{ \begin{small}
\begin{tabular}{ll}                              
Alison M. Etheridge 
 & Thomas G. Kurtz \\   
Department of Statistics & Departments of Mathematics and Statistics\\       
Oxford University & University of Wisconsin - Madison \\                   
24-29 St Giles & 480 Lincoln Drive\\                                                         
Oxford OX1 3LB & Madison, WI  53706-1388\\
UK & USA \\                        
etheridg@stats.ox.ac.uk & kurtz@math.wisc.edu     \\
\url{http://www.stats.ox.ac.uk/~etheridg/} & 
\url{http://www.math.wisc.edu/~kurtz/}  \\       \\
\\
Ian Letter&  Peter L. Ralph 
\\   
Department of Statistics & Departments of Mathematics and Biology\\
Oxford University &University of Oregon\\                   
24-29 St Giles & Fenton Hall\\
Oxford OX1 3LB & Eugene, OR 97403-1222\\
UK & USA \\
restucci@stats.ox.ac.uk  & plr@oregon.edu \\
\url{https://www.stats.ox.ac.uk/~restucci/}&
\url{https://math.uoregon.edu/profile/plr} \\
\\
Terence Tsui Ho Lung
 &  \\   
Department of Statistics & \\
Oxford University & \\                   
24-29 St Giles& \\
Oxford OX1 3LB & \\
UK & \\
terence.tsui@sjc.ox.ac.uk &      \\
\url{https://www.maths.ox.ac.uk/people/terence.tsui}&  \\
\end{tabular}
\end{small}}

\date{\today}
\maketitle

\begin{abstract}

We introduce a broad class of mechanistic spatial models to describe how spatially 
heterogeneous populations live, die, and reproduce.
Individuals are represented by points of a point measure,
whose birth and death rates can depend 
both on spatial position and local population density, defined at a location
to be the convolution of the point measure
with a suitable non-negative integrable kernel
centred on that location. We pass to three different scaling limits:
an interacting superprocess, a nonlocal partial differential equation (PDE), 
and a classical PDE. 
The classical PDE is obtained both by a two-step convergence argument, in which we 
first scale time and population size and
pass to the nonlocal PDE, 
and then scale the kernel that 
determines local population density; and 
in the important special case in which the limit is a 
reaction-diffusion equation, directly 
by simultaneously scaling the kernel width,
timescale and population size in our individual based model.

A novelty of our model is that we explicitly model a juvenile phase.
The number of juveniles produced by an individual depends
on local population density at the location of the parent;
these juvenile offspring are 
thrown off in a (possibly heterogeneous, anisotropic) Gaussian distribution 
around the location of the parent; they then reach (instant) maturity with a probability
that can depend on the local population density
% (possibly measured through convolution with a different kernel from that used in determining fecundity) 
at the location at which they land. 
Although we only record mature individuals, a trace of this two-step description 
remains in our population models, resulting in novel limits in which the spatial
dynamics are governed by a nonlinear diffusion.

Using a lookdown representation, we are able to retain information about 
genealogies relating individuals in our population and, in the case of 
deterministic limiting models, we use this to deduce 
the backwards in time motion of the ancestral lineage of an individual
sampled from the population. We observe that
knowing the history of the population density is not enough to determine the motion of
ancestral lineages in our model. We also investigate (and
contrast) the behaviour of lineages for 
three different deterministic models of a population expanding its range as a 
travelling wave: the Fisher-KPP equation, the Allen-Cahn equation, and a porous 
medium equation with logistic growth.

\vspace{.1in}

\noindent {\bf Key words:}  population model, interacting superprocess, 
lookdown construction, porous medium equation, 
reaction-diffusion equation, travelling waves, genealogies,
Fisher-KPP equation

\vspace{.1in}

\noindent {\bf MSC 20}10 {\bf Subject Classification:}  Primary:  
%60J25, 92D10, 92D15, 92D25 92D40  
\\Secondary:   %60F05, 60G09, 60G55, 60G57, 60H15, 60J68

\end{abstract}
\tableofcontents
\newpage

@@

The history of a natural population is often only accessible indirectly, 
through patterns of genetic diversity that have been laid down; from 
genetic data, one can try to infer the genealogical trees that relate 
individuals in a sample from the population, and these have been shaped 
by its history \citep[see e.g.,][]{neigel1993application,kelleher2019inferring}.
It is therefore of interest to establish information about the distribution of genealogical trees 
under our population model, which we do with a lookdown construction. 
Lookdown constructions were first introduced in~\cite{donnelly/kurtz:1996} to provide 
a mechanism for retaining information about genealogical relationships between individuals
sampled from a population evolving according to the Moran model when passing to the 
infinite population limit.
Since then, they have been extended to a wide range of models. Of particular
relevance to our work here are the papers of~\cite{kurtz/rodrigues:2011} 
and~\cite{etheridge/kurtz:2019}, in which
lookdown constructions are provided for a wide variety of population models, including
spatially structured branching processes.

@@

\paragraph{Acknowledgements:}
Thanks go to Gilia Patterson for identifying the ``clumping'' phenomenon,
and to Marcin Bownick and David Levin for useful discussions.
AME thanks everyone in MAPS at Universit\'e Paris Cit\'e for their hospitality
during the period in which much of this research took place.
AME and PLR also thank the Kavli Institute for Theoretical Physics
for their hospitality and birdwatching opportunities;
this research was therefore supported in part by the NSF under grant \#PHY-1748958
and by the Gordon and Betty Moore Foundation grant \#2919.02.
PLR was supported by the NIH NHGRI (grant \#HG011395),
IL by the ANID/Doctorado en el extranjero doctoral scholarship, grant \#2018-72190055,
and TTHL by the EPSRC Centre for Doctoral Training in Mathematics of Random Systems: Analysis, Modelling and Simulation (EP/S023925/1)
the Deutsche Forschungsgemeinschaft under Germany's Excellence Strategy, EXC-2047/1-390685813,
the Rhodes Trust and St.~John's College, Oxford.

@@

\begin{remark}
Although this model allows fairly general birth and death mechanisms,
there are a number of limitations.
Perhaps most obviously, to simplify the notation
individuals give birth to only one offspring at a time,
although this restriction could be easily lifted
\citep[as in Section 3.4 of][]{etheridge/kurtz:2019}.
Furthermore, individuals do not move during their lifetime,
and the age of an individual does not affect its fecundity or death rate.
Finally, there is no notion of mating
(although limitations on reproduction due to availability of mates can be incorporated into the birth rate, $\gamma$),
so the lineages we follow will be uniparental.
For these reasons, the model is most obviously applicable to bacterial populations or selfing plants,
although we do not anticipate that incorporation of these complications
will change the general picture.
\end{remark}

@@

\begin{terminology*}
In our prelimiting model,
the population is represented by a point measure in which each
individual is assigned a mass $1/N$. We use the term ``population density'' for
this process, as it is supposed to measure population size relative to a 
nominal occupancy of $N$ individuals per unit area. There is no implication 
that the measure representing the population is absolutely continuous with
respect to Lebesgue measure; indeed in the prelimit it is certainly not.
\end{terminology*}

@@

\begin{figure}
    \begin{center}
        % \includegraphics{figures/ex1a/fkpp_123.density.3.0}
        % \includegraphics{figures/ex1b/fkpp_123.density.3.0}
        \includegraphics{figures/fig1a}
        \includegraphics{figures/fig1b}
    \end{center}
    \caption{
        Snapshots of two simulations, with small $\alpha=\theta/N$ (left) and large 
	$\alpha =\theta/N$ (right).
        Simulations are run with a Fisher-KPP-like parameterization:
        birth and establishment are constant, while death increases linearly with density,
        at slope $1/\theta$.
        Left: $\alpha=0.1$. Right: $\alpha=10$.
        Other parameters were the same:
        dispersal and interactions distance were set to 1,
        and the equilibrium density is 10 individuals per unit area.
        \label{fig:super_vs_det_2d}
    }
\end{figure}

@@

\begin{figure}
    \begin{center}
        % \includegraphics[width=0.7\textwidth]{figures/ex2a/pme_1234.steps}
        % \includegraphics[width=0.7\textwidth]{figures/ex2b/pme_1234.steps}
        \includegraphics[width=0.7\textwidth]{figures/fig2a}
        \includegraphics[width=0.7\textwidth]{figures/fig2b}
    \end{center}
    \caption{
        Simulated populations under a porous medium equation with logistic growth~(\ref{PME}) in $d=1$,
        % comparing the simulated profile to the analytical solution;
        $\theta/N$ small on the left; large on the right.
        % Simulation is with $\gamma(m) = m$ and $\mu(m) = (1+1/\theta) m - 1/\theta$
        % (implemented by choosing a small $dt$ taking reproduction probability per time step
        % to be $1 - \exp(- dt m)$
        % and death probability to be $\mu(m) = 1 - \exp(-dt ((1+1/\theta) m - 1/\theta))$);
        Values of $\theta$ in top and bottom figures
        are 1 and 100, respectively, and both have $N$ set
        so that the density is roughly 100 individuals per unit of habitat
        (as displayed on the vertical axis).
        See text for details of the simulations.
        \label{fig:pme_waves}
    }
\end{figure}

@@

\begin{figure}
    \begin{center}
        % \includegraphics[width=0.45\textwidth]{figures/clumpy_2d/clumpy_123.density.3.0}
        % \includegraphics[width=0.45\textwidth]{figures/clumpy_1d/clumpy_123.steps}
        \includegraphics[width=0.45\textwidth]{figures/fig3a}
        \includegraphics[width=0.45\textwidth]{figures/fig3b}
    \end{center}
    \caption{
        \textbf{Left:} A snapshot of individual locations in a two-dimensional simulation
        in which the constant density is unstable
        and a stable, periodic pattern forms.
        \textbf{Right:} Population density in an expanding wave in a one-dimensional simulation
        forming a periodic pattern;
        each panel shows the wavefront in three periods of time;
        within each period of time the wavefront at earlier times is shown in blue
        and later times in pink.
        In both cases, $\gamma(m) = 3/(1 + m)$, $\mu \equiv 0.3$, and $r \equiv 1$;
        dispersal is Gaussian with $\sigma=0.2$ and density is measured with $\rho_\gamma(x) = p_9(x)$,
        i.e., using a Gaussian kernel with standard deviation 3.
        \label{fig:clumping}
    }
\end{figure}

@@

\bibliographystyle{plainnat}
\bibliography{refs}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                               %%
%% This is the sample.tex file for the ejpecp document class.    %%
%% This file is for ejpecp version 1.0                           %%
%% Please be sure that you are using the lastest version:        %%
%% https://www.ctan.org/pkg/ejpecp                               %%
%%                                                               %%
%% The ejpecp class works *only* with a pdflatex engine.         %%
%% You need the ejpecp.cls in your current directory or in any   %%
%% directory scanned for cls files by your pdflatex engine.      %%
%%                                                               %%
%% Manual inclusion of page layout commands is useless.          %%
%%                                                               %%
%% Note that any complex file will produce delayed publication!  %%
%%                                                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                               %%
%% Journal selection: ECP or EJP.                                %%
%%                                                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[EJP]{ejpecp} % replace ECP by EJP if needed.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                               %%
%% Please uncomment and adapt to your encoding if needed:        %%
%%                                                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\usepackage[T1]{fontenc}
%\usepackage[utf8]{inputenc}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                               %%
%% Please add here your own packages (be minimalistic please!):  %%
%% Please avoid using exotic packages and keep things simple.    %%
%% It is not necessary to include ams* and graphicx packages     %%
%% since they are automatically included by the ejpecp class.    %%
%%                                                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\usepackage{enumerate}  % uncomment to use this package

% \usepackage{xcolor}
% \usepackage[hidelinks]{hyperref}
% \usepackage{bbm}
% \usepackage{natbib}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                               %%
%% Shorttitle (please edit and customize for running heading):   %%
%% Title (please edit and customize):                            %%
%%                                                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\SHORTTITLE{
Dynamics and genealogies of locally regulated populations
}

\TITLE{
Looking forwards and backwards: dynamics and genealogies of locally regulated populations
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                               %%
%% Authors (please edit and customize):                          %%
%%                                                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\AUTHORS{%
    Alison~M.~Etheridge\footnote{Oxford University, UK}
    \footnote{\EMAIL{etheridge@stats.ox.ac.uk}}
  \and
  Thomas~G.~Kurtz\footnote{University of Wisconsin - Madison, USA:
    \EMAIL{kurtz@math.wisc.edu}}
  \and
  Ian~Letter\footnotemark[1]
  \and
  Peter~L.~Ralph\footnote{University of Oregon, USA:
    \EMAIL{plr@uoregon.edu}}
  \and
  Terence~Tsui~Ho~Lung\footnotemark[1]
}%AUTHORS

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                               %%
%% Please edit and customize the following items:                %%
%%                                                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\KEYWORDS{population model ; interacting superprocess ; lookdown construction ; porous medium equation ; reaction-diffusion equation ; travelling waves ; genealogies ; Fisher-KPP equation}
\AMSSUBJ{60J25 ; 92D10 ; 92D15 ; 92D25 ; 92D40}
\AMSSUBJSECONDARY{60F05 ; 60G09 ; 60G55 ; 60G57 ; 60H15 ; 60J68}

\SUBMITTED{June 14, 2023} % Edit.
\ACCEPTED{NA} % Edit.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                               %%
%% Please uncomment and edit if you have an arXiv ID:            %%
%%                                                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\ARXIVID{2305.14488}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                               %%
%% The following items will be set by the Managing Editor.       %%
%%                                                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\VOLUME{0}
\YEAR{2020}
\PAPERNUM{0}
\DOI{10.1214/YY-TN}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                               %%
%% Please edit and customize the abstract:                       %%
%%                                                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\ABSTRACT{
We introduce a broad class of mechanistic spatial models to describe how spatially heterogeneous populations live, die, and reproduce.  Individuals are represented by points of a point measure, whose birth and death rates can depend both on spatial position and local population density, defined at a location to be the convolution of the point measure with a suitable non-negative integrable kernel centred on that location. We pass to three different scaling limits: an interacting superprocess, a nonlocal partial differential equation (PDE), and a classical PDE.  The classical PDE is obtained both by a two-step convergence argument, in which we first scale time and population size and pass to the nonlocal PDE, and then scale the kernel that determines local population density; and in the important special case in which the limit is a reaction-diffusion equation, directly by simultaneously scaling the kernel width, timescale and population size in our individual based model.

A novelty of our model is that we explicitly model a juvenile phase.  The number of juveniles produced by an individual depends on local population density at the location of the parent; these juvenile offspring are thrown off in a (possibly heterogeneous, anisotropic) Gaussian distribution around the location of the parent; they then reach (instant) maturity with a probability that can depend on the local population density at the location at which they land.  Although we only record mature individuals, a trace of this two-step description remains in our population models, resulting in novel limits in which the spatial dynamics are governed by a nonlinear diffusion.

Using a lookdown representation, we are able to retain information about genealogies relating individuals in our population and, in the case of deterministic limiting models, we use this to deduce the backwards in time motion of the ancestral lineage of an individual sampled from the population. We observe that knowing the history of the population density is not enough to determine the motion of ancestral lineages in our model. We also investigate (and contrast) the behaviour of lineages for three different deterministic models of a population expanding its range as a travelling wave: the Fisher-KPP equation, the Allen-Cahn equation, and a porous medium equation with logistic growth.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                               %%
%% Please add your own macros and environments below:            %%
%%                                                               %%
%% If possible, avoid using \def and use instead \newcommand     %%
%% If possible, avoid defining your own environments, and use    %%
%% instead the environments already defined by ejpecp:           %%
%%  assumption, assumptions, claim, condition, conjecture,       %%
%%  corollary, definition, definitions, example, exercise, fact, %%
%%  facts, heuristics, hypothesis, hypotheses, lemma, notation,  %%
%%  notations, problem, proposition, remark, theorem             %%
%%                                                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\renewcommand{\theequation}{\arabic{section}.\arabic{equation}}
\renewcommand{\hat}{\widehat}
\renewcommand{\tilde}{\widetilde}
\renewcommand{\bar}{\overline}
\newcommand{\IP}{\mathbb P}
\newcommand{\IQ}{\mathbb Q}
\newcommand{\IE}{\mathbb E}
\newcommand{\IR}{\mathbb R}
\newcommand{\IZ}{\mathbb Z}
\newcommand{\IN}{\mathbb N}
\newcommand{\IT}{\mathbb T}
\newcommand{\IC}{\mathbb C}
\newcommand{\ind}{\mathbf{1}}
\newcommand{\bigO}{\mathcal{O}}
\newcommand{\grad}{\nabla}
\newcommand{\dif}{\mathrm{d}\,}

\newcommand{\DG}{\mathcal{B}}  % generator of the dispersal process
\newcommand{\DD}{\mathcal{D}}  % the second-order part of the generator of dispersal
\newcommand{\meanq}{\vec b}    % mean of dispersal, times theta
\newcommand{\covq}{\mathbf{C}}     % covariance matrix of dispersal, times theta
\newcommand{\kernel}{\rho}  % interaction kernels
\newcommand{\smooth}[1]{\kernel_{#1} \! * \!}  % convolution by the interaction kernel
\newcommand{\wavespeed}{\mathfrak{c}}    % speed (vector) of a wave
\newcommand{\Lgen}{\mathcal{L}}    % generator of a lineage
\newcommand{\Pgen}{\mathcal{P}}    % generator of the population process
\newcommand{\lp}{\xi}              % process with levels
\newcommand{\labelspace}{\mathcal{I}} % space of labels
\newcommand{\concat}{\oplus}   % concatenation of labels
\newcommand{\measures}{\mathcal{M}_F(\IR^d)} % finite measures on Rd
\newcommand{\cmeasures}{\mathcal{M}_F(\overline{\IR}^d)} % finite measures on compactified Rd
\newcommand{\lpmeasures}{\mathcal{M}(\overline{\IR^d} \times [0,\infty))} % locally finite measures on (compactified space) x level
\newcommand{\flpmeasures}{\mathcal{M}(\IR^d \times [0,\infty))} % locally finite measures on space x level

\newcommand{\citet}[1]{\cite{#1}}
\newcommand{\citep}[1]{\cite{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                               %%
%% No macro definitions below this line please!                  %%
%%                                                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                               %%
%% No need for \maketitle.                                       %%
%%                                                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \author{ \begin{small}
% \begin{tabular}{ll}                              
% Alison M. Etheridge 
%  & Thomas G. Kurtz \\   
% Department of Statistics & Departments of Mathematics and Statistics\\       
% Oxford University & University of Wisconsin - Madison \\                   
% 24-29 St Giles & 480 Lincoln Drive\\                                                         
% Oxford OX1 3LB & Madison, WI  53706-1388\\
% UK & USA \\                        
% etheridg@stats.ox.ac.uk & kurtz@math.wisc.edu     \\
% \url{http://www.stats.ox.ac.uk/~etheridg/} & 
% \url{http://www.math.wisc.edu/~kurtz/}  \\       \\
% \\
% Ian Letter&  Peter L. Ralph 
% \\   
% Department of Statistics & Departments of Mathematics and Biology\\
% Oxford University &University of Oregon\\                   
% 24-29 St Giles & Fenton Hall\\
% Oxford OX1 3LB & Eugene, OR 97403-1222\\
% UK & USA \\
% restucci@stats.ox.ac.uk  & plr@oregon.edu \\
% \url{https://www.stats.ox.ac.uk/~restucci/}&
% \url{https://math.uoregon.edu/profile/plr} \\
% \\
% Terence Tsui Ho Lung
%  &  \\   
% Department of Statistics & \\
% Oxford University & \\                   
% 24-29 St Giles& \\
% Oxford OX1 3LB & \\
% UK & \\
% terence.tsui@sjc.ox.ac.uk &      \\
% \url{https://www.maths.ox.ac.uk/people/terence.tsui}&  \\
% \end{tabular}
% \end{small}}

@@

The history of a natural population is often only accessible indirectly, 
through patterns of genetic diversity that have been laid down; from 
genetic data, one can try to infer the genealogical trees that relate 
individuals in a sample from the population, and these have been shaped 
by its history (see e.g., \citep{neigel1993application,kelleher2019inferring}).
It is therefore of interest to establish information about the distribution of genealogical trees 
under our population model, which we do with a lookdown construction. 
Lookdown constructions were first introduced in~\cite{donnelly/kurtz:1996} to provide 
a mechanism for retaining information about genealogical relationships between individuals
sampled from a population evolving according to the Moran model when passing to the 
infinite population limit.
Since then, they have been extended to a wide range of models. Of particular
relevance to our work here are the papers of~\cite{kurtz/rodrigues:2011} 
and~\cite{etheridge/kurtz:2019}, in which
lookdown constructions are provided for a wide variety of population models, including
spatially structured branching processes.

@@

@@

\begin{remark}
Although this model allows fairly general birth and death mechanisms,
there are a number of limitations.
Perhaps most obviously, to simplify the notation
individuals give birth to only one offspring at a time,
although this restriction could be easily lifted
(as in Section 3.4 of \cite{etheridge/kurtz:2019}).
Furthermore, individuals do not move during their lifetime,
and the age of an individual does not affect its fecundity or death rate.
Finally, there is no notion of mating
(although limitations on reproduction due to availability of mates can be incorporated into the birth rate, $\gamma$),
so the lineages we follow will be uniparental.
For these reasons, the model is most obviously applicable to bacterial populations or selfing plants,
although we do not anticipate that incorporation of these complications
will change the general picture.
\end{remark}

@@

\begin{remark}
In our prelimiting model,
the population is represented by a point measure in which each
individual is assigned a mass $1/N$. We use the term ``population density'' for
this process, as it is supposed to measure population size relative to a 
nominal occupancy of $N$ individuals per unit area. There is no implication 
that the measure representing the population is absolutely continuous with
respect to Lebesgue measure; indeed in the prelimit it is certainly not.
\end{remark}

@@

\begin{figure}
    \begin{center}
        % \includegraphics{figures/ex1a/fkpp_123.density.3.0}
        % \includegraphics{figures/ex1b/fkpp_123.density.3.0}
        \includegraphics[width=2.5in]{figures/fig1a}
        \includegraphics[width=2.5in]{figures/fig1b}
    \end{center}
    \caption{
        Snapshots of two simulations, with small $\alpha=\theta/N$ (left) and large 
	$\alpha =\theta/N$ (right).
        Simulations are run with a Fisher-KPP-like parameterization:
        birth and establishment are constant, while death increases linearly with density,
        at slope $1/\theta$.
        Left: $\alpha=0.1$. Right: $\alpha=10$.
        Other parameters were the same:
        dispersal and interactions distance were set to 1,
        and the equilibrium density is 10 individuals per unit area.
        \label{fig:super_vs_det_2d}
    }
\end{figure}

@@

\begin{figure}
    \begin{center}
        % \includegraphics[width=0.7\textwidth]{figures/ex2a/pme_1234.steps}
        % \includegraphics[width=0.7\textwidth]{figures/ex2b/pme_1234.steps}
        \includegraphics[width=4.5in]{figures/fig2a}
        \includegraphics[width=4.5in]{figures/fig2b}
    \end{center}
    \caption{
        Simulated populations under a porous medium equation with logistic growth~(\ref{PME}) in $d=1$,
        % comparing the simulated profile to the analytical solution;
        $\theta/N$ small on the left; large on the right.
        % Simulation is with $\gamma(m) = m$ and $\mu(m) = (1+1/\theta) m - 1/\theta$
        % (implemented by choosing a small $dt$ taking reproduction probability per time step
        % to be $1 - \exp(- dt m)$
        % and death probability to be $\mu(m) = 1 - \exp(-dt ((1+1/\theta) m - 1/\theta))$);
        Values of $\theta$ in top and bottom figures
        are 1 and 100, respectively, and both have $N$ set
        so that the density is roughly 100 individuals per unit of habitat
        (as displayed on the vertical axis).
        See text for details of the simulations.
        \label{fig:pme_waves}
    }
\end{figure}

@@

\begin{figure}
    \begin{center}
        % \includegraphics[width=0.45\textwidth]{figures/clumpy_2d/clumpy_123.density.3.0}
        % \includegraphics[width=0.45\textwidth]{figures/clumpy_1d/clumpy_123.steps}
        \includegraphics[width=2.5in]{figures/fig3a}
        \includegraphics[width=2.5in]{figures/fig3b}
    \end{center}
    \caption{
        \textbf{Left:} A snapshot of individual locations in a two-dimensional simulation
        in which the constant density is unstable
        and a stable, periodic pattern forms.
        \textbf{Right:} Population density in an expanding wave in a one-dimensional simulation
        forming a periodic pattern;
        each panel shows the wavefront in three periods of time;
        within each period of time the wavefront at earlier times is shown in blue
        and later times in pink.
        In both cases, $\gamma(m) = 3/(1 + m)$, $\mu \equiv 0.3$, and $r \equiv 1$;
        dispersal is Gaussian with $\sigma=0.2$ and density is measured with $\rho_\gamma(x) = p_9(x)$,
        i.e., using a Gaussian kernel with standard deviation 3.
        \label{fig:clumping}
    }
\end{figure}

@@

% \bibliographystyle{amsplain}
% \bibliography{refs}

% add below the content of your .bbl file produced by bibtex.

\providecommand{\bysame}{\leavevmode\hbox to3em{\hrulefill}\thinspace}
\providecommand{\MR}{\relax\ifhmode\unskip\space\fi MR }
% \MRhref is called by the amsart/book/proc definition of \MR.
\providecommand{\MRhref}[2]{%
  \href{http://www.ams.org/mathscinet-getitem?mr=#1}{#2}
}
\providecommand{\href}[2]{#2}
\begin{thebibliography}{10}

\bibitem{barlow/jacka/yor:1986}
M.~T. Barlow, S.~D. Jacka, and M.~Yor, \emph{Inequalities for a pair of
  processes stopped at a random time}, Proceedings of the London Mathematical
  Society \textbf{s3-52} (1986), no.~1, 142--172.

\bibitem{barton:1979}
N~H Barton, \emph{The dynamics of hybrid zones}, Heredity \textbf{43} (1979),
  no.~3, 341--359.

\bibitem{berestycki2009nonlocal}
Henri Berestycki, Gregoire Nadin, Benoit Perthame, and Lenya Ryzhik, \emph{The
  non-local fisher-kpp equation: travelling waves and steady states},
  Nonlinearity \textbf{22} (2009), no.~12, 2813--2844.

\bibitem{birzu/hallatschek/korolev:2017}
Gabriel Birzu, Oskar Hallatschek, and Kirill~S. Korolev, \emph{Fluctuations
  uncover a distinct class of traveling waves}, Proc. Natl. Acad. Sci. (USA)
  \textbf{115} (2018), no.~16, 3645--3654.

\bibitem{birzu/hallatschek/korolev:2021}
Gabriel Birzu, Oskar Hallatschek, and Kirill~S Korolev, \emph{Genealogical
  structure changes as range expansions transition from pushed to pulled},
  Proc. Natl. Acad. Sci. (USA) \textbf{118} (2021), no.~34, e2026746118.

\bibitem{biswas/etheridge/klimek:2018}
Niloy Biswas, Alison Etheridge, and Aleksander Klimek, \emph{The spatial
  {Lambda}-{Fleming}-{Viot} process with fluctuating selection}, Electron. J.
  Probab. \textbf{26} (2021), 1 -- 51.

\bibitem{bolker/pacala:1997}
B~M Bolker and S~W Pacala, \emph{Using moment equations to understand
  stochastically driven spatial pattern formation in ecological systems},
  Theor. Pop. Biol. \textbf{52} (1997), no.~3, 179--197.

\bibitem{bolker/pacala:1999}
\bysame, \emph{Spatial moment equations for plant competition: {Understanding}
  spatial strategies and the advantages of short dispersal}, American
  Naturalist \textbf{153} (1999), no.~6, 575--602.

\bibitem{bramson:1983}
M~Bramson, \emph{{Convergence of solutions of the Kolmogorov equation to
  travelling waves}}, Mem. Amer. Math. Soc. \textbf{44} (1983), no.~285.

\bibitem{britton1990spatial}
N.~F. Britton, \emph{Spatial structures and periodic travelling waves in an
  integro-differential reaction-diffusion population model}, SIAM Journal on
  Applied Mathematics \textbf{50} (1990), no.~6, 1663--1688.
